---
title: "US Police Shootings"
author: "Ketan Mittal"
date: "`r Sys.Date()`"
output: html_document
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

```{r}
shooting_data <- read_csv("shootings.csv")
attach(shooting_data)
```
```{r}
print(sum(shooting_data$gender == 'F')/nrow(shooting_data))
print(sum(shooting_data$signs_of_mental_illness == 'TRUE')/nrow(shooting_data))
print(sum(shooting_data$body_camera == 'TRUE')/nrow(shooting_data))
```
Around 95% of the people shot by the police are male and around 5% of the people shot were female. Around 23% of people killed had a mental illness while 77% didn't. Around 12% of the cases had body cameras involved while 88% didn't. 

```{r}
number_ticks <- function(n) {function(limits) pretty(limits, n)}
ggplot(shooting_data, aes(x = factor(age), fill = race)) + 
  geom_histogram(aes(y = ..count..), stat = "count") +
  scale_x_discrete(breaks = number_ticks(25)) +
  labs(title = "People's Age when Killed", y = "percentage", fill = "Race", x = "Age")
```
As you can see police killings usually occurs from ages 25 to 40. 

```{r}
state_data <- 
  shooting_data %>% 
  mutate(state = case_when(
    state == "CA" | state == "NV" |state == "AK" | state == "HI" ~ "West", 
    state == "OR"| state == "WA"| state ==  "ID"| state ==  "MT"| state ==  "WY" ~ "North West", 
    state == "UT"| state ==  "AZ"| state ==  "NM"| state ==  "CO"| state ==  "TX"| state ==  "OK" ~ "South West", 
    state == "ND"| state ==  "SD"| state ==  "NE"| state ==  "KS"| state ==  "MN"| state ==  "IA"| state ==  "MO"| state ==  "WI"| state ==  "IL"| state ==  "IN"| state ==  "MI"| state ==  "OH"| state ==  "KY" ~ "Mid-West",
    state == "AR"| state ==  "LA"| state ==  "TN"| state ==  "MS"| state ==  "AL"| state ==  "GA"| state ==  "FL"| state ==  "SC"| state ==  "NC" ~ "South East",
    state == "WV"| state ==  "VA"| state ==  "MD"| state ==  "DE"| state ==  "PA"| state ==  "NY"| state ==  "NJ"| state == "DC" ~ "Mid-Atlantic",
    state == "VT"| state ==  "CT"| state ==  "MA"| state ==  "NH"| state ==  "ME"| state ==  "RI" ~ "North East"
  ))
ggplot(state_data, aes(x = factor(state), fill = race)) + 
  geom_bar(aes(y = ((..count..)/sum(..count..))), stat = "count") +
  scale_y_continuous(labels=scales::percent) +
  labs(title = "The Number of Police Killings by State Region described by Race", y = "percentage", fill = "Race", x = "State Regions")
       
```
We can see the distrubitions of police shootings per region and the distribution of race within that. 

```{r}
ggplot(state_data, aes(x = factor(race), fill = race)) + 
  geom_bar(aes(y = ((..count..)/sum(..count..))), stat = "count") +
  geom_text(aes(label = scales::percent(round((..count..)/sum(..count..), 2)),
                y = ((..count..)/sum(..count..))), stat = "count", vjust = -.25) + 
  scale_y_continuous(labels=scales::percent) + 
  labs(title = "Number of Police Killings by Race in Percent", y = "percentage", fill = "Race", x = "Race")
```
We can see the police killings by Race. 
```{r}
gov_site <- data.frame(
    race = c("Asian", "Black", "Hispanic", "Native", "Other", "White"),
    percentages = c(6.1, 13.6, 18.9, 1.3, .8, 59.3)
    )
ggplot(gov_site, aes(x = factor(race), fill = race)) + 
  geom_bar(aes(y = (percentages/100)), stat = "Identity") +
  geom_text(aes(label = scales::percent(percentages/100),
                y = percentages/100), stat = "Identity", vjust = -.25) + 
  scale_y_continuous(labels=scales::percent) + 
  labs(title = "Percent of People in America", y = "percentage", fill = "Race", x = "Race")
```
We can see the percent of races in America. 
We can see how certain races have a higher or lower chance of being killed by the police.

Since the Montana 1948 focuses on racism on Native Americans, here is my data analysis on it. 
```{r}
native_data <- 
  subset(state_data, race=="Native")
native_data <- 
  native_data %>% 
  mutate(arms_category = case_when(
    arms_category == "Unarmed" ~"Unarmed",
    arms_category != "Unarmed" ~ "Weapon or Unknown"
  ))
ggplot(native_data, aes(x = arms_category, fill = flee)) + 
  geom_bar(aes(y = ((..count..)/sum(..count..))), stat = "count") +
  scale_y_continuous(labels=scales::percent) + 
  labs(title = "Percent of Native Americans that were Killed based on whether the person \n was Armed or Fleeing", y = "percentage", fill = "How the Person Fled", x = "Armed or Not")
```
As you can see most of the police shootings against Native Americans were justified, however, there were a few that weren't. 

```{r}
years <- as.character(format(shooting_data$date, "%Y"))
yValues <- c(sum(years == "2015"), sum(years == "2016"), sum(years == "2017"), sum(years == "2018"), sum(years == "2019"))
xValues <- 0:4
date <- data.frame(xValues, yValues)
ggplot(date, aes(x=xValues, y=yValues)) +
  geom_line(color = "steelblue", size=2, alpha=0.75, linetype=7) + 
  labs(title = "Number of Police Killings from 2015 to 2019", x = "Years from 2015", y = "Number of Police Killings")
```
From 2015 to 2019 the number of police killings decrease. 
